---
title: "Analysis on Youtube data (Stat 154 Final Project)"
author: "Danielle Ip, Youngshin Kim"
output:
  html_document:
    df_print: paged
---
Description of the dataset

```{r}
#Reading dataset 
youtube = read.csv("USvideos.csv", stringsAsFactors = FALSE, header = TRUE)

#Libraries
library(lubridate)
library(randomForest)
library(glmnet)
library(dplyr)
library(tree)
```

```{r}
#summary(youtube)
#head(youtube)
```



```{r}

#Adding new column "days_to_trending" which is the number of days 
#a video took to go trending
youtube$trending_date = as.Date(youtube$trending_date, "%y/%d/%m")
youtube$publish_time = as.Date(as.character(substr(youtube$publish_time, start = 1,stop = 10)))
youtube$days_to_trending = youtube$trending_date - youtube$publish_time
lct <- Sys.getlocale("LC_TIME"); Sys.setlocale("LC_TIME", "C")
Sys.setlocale("LC_TIME", lct)
summary(youtube)
#youtube[!complete.cases(youtube),]
```

```{r}
#Adding new column "month" which is the month the video was published
youtube$month_num = month(youtube$publish_time)
youtube$month_num = as.numeric(youtube$month_num)
youtube[youtube$month_num == 1,"month"] = 'January'
youtube[youtube$month_num == 2,"month"] = 'February'
youtube[youtube$month_num == 3,"month"] = 'March'
youtube[youtube$month_num == 4,"month"] = 'April'
youtube[youtube$month_num == 5,"month"] = 'May'
youtube[youtube$month_num == 6,"month"] = 'June'
youtube[youtube$month_num == 7,"month"] = 'July'
youtube[youtube$month_num == 8,"month"] = 'August'
youtube[youtube$month_num == 9,"month"] = 'September'
youtube[youtube$month_num == 10,"month"] = 'October'
youtube[youtube$month_num == 11,"month"] = 'November'
youtube[youtube$month_num == 12,"month"] = 'December'

#youtube$publish_time = NULL
#youtube$trending_date = NULL
youtube$month_num = NULL

#Changing category_id from numbers to actual categories
youtube[youtube$category_id == 1,"category"] = 'Film and Animation'
youtube[youtube$category_id == 2,"category"] = 'Cars and Vehicles'
youtube[youtube$category_id == 10,"category"] = 'Music'
youtube[youtube$category_id == 15,"category"] = 'Pets and Animals'
youtube[youtube$category_id == 17,"category"] = 'Sport'
youtube[youtube$category_id == 18,"category"] = 'Short Movies'
youtube[youtube$category_id == 19,"category"] = 'Travel and Events'
youtube[youtube$category_id == 20,"category"] = 'Gaming'
youtube[youtube$category_id == 21,"category"] = 'Videoblogging'
youtube[youtube$category_id == 22,"category"] = 'People and Blogs'
youtube[youtube$category_id == 23,"category"] = 'Comedy'
youtube[youtube$category_id == 24,"category"] = 'Entertainment'
youtube[youtube$category_id == 25,"category"] = 'News and Politics'
youtube[youtube$category_id == 26,"category"] = 'How to and Style'
youtube[youtube$category_id == 27,"category"] = 'Education'
youtube[youtube$category_id == 28,"category"] = 'Science and Technology'
youtube[youtube$category_id == 29,"category"] = 'Non Profits and Activism'
youtube[youtube$category_id == 30,"category"] = 'Movies'
youtube[youtube$category_id == 31,"category"] = 'Anime/Animation'
youtube[youtube$category_id == 32,"category"] = 'Action/Adventure'
youtube[youtube$category_id == 33,"category"] = 'Classics'
youtube[youtube$category_id == 34,"category"] = 'Comedy'
youtube[youtube$category_id == 35,"category"] = 'Documentary'
youtube[youtube$category_id == 36,"category"] = 'Drama'
youtube[youtube$category_id == 37,"category"] = 'Family'
youtube[youtube$category_id == 38,"category"] = 'Foreign'
youtube[youtube$category_id == 39,"category"] = 'Horror'
youtube[youtube$category_id == 40,"category"] = 'Sci-Fi/Fantasy'
youtube[youtube$category_id == 41,"category"] = 'Thriller'
youtube[youtube$category_id == 42,"category"] = 'Shorts'
youtube[youtube$category_id == 43,"category"] = 'Shows'
youtube[youtube$category_id == 44,"category"] = 'Trailers'

youtube$category_id = NULL

#Make values numeric
youtube$views = as.numeric(youtube$views)
youtube$likes = as.numeric(youtube$likes)
youtube$dislikes = as.numeric(youtube$dislikes)
youtube$comment_count = as.numeric(youtube$comment_count)
youtube$days_to_trending = as.numeric(youtube$days_to_trending)
#youtube$video_id <- as.factor(youtube$video_id)
#youtube$title <- as.factor(youtube$title)
youtube$channel_title <- as.factor(youtube$channel_title)
#youtube$thumbnail_link <- as.factor(youtube$thumbnail_link)
youtube$category <- as.factor(youtube$category)
#youtube$description <- as.factor(youtube$description)
monthOrder <- c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")
youtube$month <- as.factor(youtube$month)
youtube$month <- factor(youtube$month, levels=monthOrder)



head(youtube)
summary(youtube)
```





EDA
```{r}
#plot(table(youtube$category))
#plot(table(youtube$month))
library(ggplot2)
ggplot(youtube, aes(x=views)) + geom_histogram() + xlim(0, 2000000) + labs(title="Number of Views Distribution")
ggplot(youtube, aes(x=likes)) + geom_histogram() + xlim(0, 60000) + labs(title="Number of Likes Distribution")
ggplot(youtube, aes(month, days_to_trending)) + geom_boxplot() + labs(title="Publish month versus days to trending")
ggplot(youtube, aes(x=days_to_trending)) + geom_bar() +xlim(0, 25) + labs(title="Days to Trending Distribution")

```

Logisitic Regression 
```{r}
youtube$fivedays <- youtube$days_to_trending <= 5
youtube$waitmonth <- youtube$month %in% c("July", "August", "September", "October")
youtube$topcategory <- youtube$category %in% c("Entertainment", "Music", "How to and Style")
#summary(youtube)
#summary(youtube[youtube$waitmonth == 1,])
set.seed(12)
train <- sample(nrow(youtube), nrow(youtube)/50)
trainset <- youtube[train, ]
testset <- youtube[-train, ]
glm.fits <- glm(fivedays ~ views + likes + dislikes  + comment_count + comments_disabled + waitmonth + topcategory, data=trainset, family=binomial)
summary(glm.fits)
#plot(glm.fits)
```

```{r}
backwards <- step(glm.fits)
```

```{r}
glm.selected <- glm(fivedays ~ views + likes + dislikes + waitmonth, data=trainset, family=binomial)
summary(glm.selected)
```
```{r}
glm.probs <- predict(glm.selected, testset, type="response")
glm.pred <- rep("0", nrow(testset))
glm.pred[glm.probs>0.5] = "1"
table(glm.pred, testset$fivedays)
```

Interpretation for Logisitic Regression Model 

The model seemed to be improved after the backwards selection process, however the finding that emphasis on likes, dislikes, and views correlated with videos that trend in five days or less is not strongly supported. The error rate is very high, this could be attributed to how all videos are very different and there are more complexities to determining a videos success beyond the KPIs available from the YouTUbe API. 


Linear Regression
```{r}
#Deleting Categorical Variables
lm.fit = lm(views~likes+dislikes+comment_count+days_to_trending, data=youtube)
summary(lm.fit)
plot(lm.fit)
```
Interpretation of the linear regression

Our R-squared value is greater than 0.78, so the model seems 
to be a good fit. However, the residuals do not seem uniformly random, therefore the data fails to meet normality assumptions. 


Random Forest
```{r}
#Splitting data into train and test data
train=sample(1:nrow(youtube), nrow(youtube)/2)
youtube.test=youtube[-train, "views"]

bag.youtube=randomForest(views~likes+dislikes+comment_count+days_to_trending, data=youtube, subset=train, mtry=4, importance=TRUE)
yhat.bag=predict(bag.youtube, newdata=youtube[-train,])
plot(yhat.bag, youtube.test)
abline(0, 1)
mean((yhat.bag-youtube.test)^2)

importance(bag.youtube)
varImpPlot(bag.youtube)
```

Interpretation of Random Forest Model

IncMSE is the MSE with the original dataset and the permuted dataset
The variables with high IncMSE would be the predictors that matter
the most in the datase.  Similarly, IncNodePurity is the increase in node purity or 
decrease in node impurity, so the variables with high node purity
values are the important predictors in the dataset.  By this definition,
it seems like likes and dislikes are important indicators of the number of views
a video will have, because they have large %IncMSE and different order of magnitude
compared to the other variables.



